Deterministic Scheduling of Periodic Messages for Cloud RAN

### 2017-07-21 infocom/infocom.tex

\begin{abstract}
A recent trend in mobile networks is to centralize the processing units which were attached to 
antennas in distant data-centers. The main challenge is to guarantee the maximum latency of the periodic messages 
sent from the antennas to their processing units and back. We show that traditional statistical multiplexing
is bad at ensuring a low latency. Hence, we propose in this article to use a deterministic scheme to send periodic messages without collisions in the network and thus without buffering and additional latency. 
We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas.
We show that there is always a solution when the routes are small or the load is small. When the parameters are unconstrained,
and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method.
The experimental results show that even under full load, most of the time PMLS finds a deterministic sending scheme with no latency.
\end{abstract}

@@

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN, or Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost ownership. The main challenge for this type of architecture is the latency in the transfer process that is an important factor for dimensioning the network. The latency is measured between the sending of a message by the Remote Radio Heads (RRHs) on the field and the receptions of the answers, computed by real-time virtualized network functions on Baseband Units (BBUs), in the cloud. For example, even if next radio protocols are less stringent, some standards requires meeting time constraints for functions like HARQ (Hybrid Automatic Repeat reQuest) that needs to be processed in $1$ to $10$ms depending on targeted services~\cite{bouguen2012lte}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer between RRHs and BBUs.

@@

      In Sec.~\ref{sec:PALL} we will study algorithms which solve the search version of \pall (computing an assignment), also denoted by \pall for simplicity. \todo{phrase utile finalement ?} 
      It turns out that the problem \pall is hard to solve for \emph{general} routed networks. The long version of this article contains complete proofs of the $\NP$-hardness or hardness of approximation of \pall and of several of its variants.

@@

       We consider a simpler version of \pall, that we call {\bf Periodic Assignment for Zero Latency} or \pazl: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce \pazl because it is simpler to study and we are able to prove theoretical results and find better algorithms than for \pall. Moreover, as we show in our experimentations of Sec.\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to \pazl is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.

       We were not able to prove that \pazl is $\NP$-hard and we leave this question open for further work. 
       It seems very likely to be hard, since a variant of \pazl where the central link is unidirectional (collisions can happen between forward and backward routes) can be shown to be $\NP$-complete. It is done by a reduction from the subset sum problem as for a similar problem of scheduling pair of tasks~\cite{orman1997complexity}.

@@

             Let us define a family of simple routed networks that we will study in this paper. 
      The graph $G$ has two sets of vertices, $S=\{s_0,...,s_{n-1}\}$ and $T=\{t_0,...,t_{n-1}\}$ of cardinality $n$ and two special nodes, the central source node $c_s$ and the central target node $c_t$.
      There is an arc between $c_s$ and  $c_t$ called the central arc and for all $i<n$ there is an arc between $s_i$ and $c_s$ and between $t_i$ and $c_t$. All the symmetric arcs are also in the graph with the same weights.
      The forward routes are the directed paths $r_i = (s_i,c_s,c_t,t_i)$ and the backward routes are $\rho(r_i) = (t_i,c_t,c_s,s_i)$. The routed network $(G, \{r_i,\rho(r_i)\}_{i<n})$ is called a \textbf{star routed network}.

@@

  Even if this topology seems really basics, it's already complex to find some solutions for \pazl and \pall on it. Also, a star routed network corresponds to a Point-to-Multipoint fronthaul (PtMP), intended for C-RAN \cite{tayq2017real}.

@@

@@

% %        We conjecture that the problem PAZL is also $\NP$-complete,  but we have yet not been able to prove it.
%   On the other hand we show positive results:  when the period is large or when the routes are short there is always a solution to PAZL and it can be found in polynomial time. We then give an exponential time exhaustive search algorithm which always finds a solution if there is one.

%      The following table summarize the main notations used in the paper.
%      \begin{center}
%   \begin{tabular}{|c|c|}
%    \hline
%     $(G,\cal R)$ & Routed network \\
%     \hline
%      $\omega(u,v)$ & Weight of the arc $(u,v) \in A$ \\
%      \hline
%      $\lambda(u_i,r)$ & Latency of the vertex $u_i$ in $r$\\
%         \hline
%         $\lambda(r)$ & Length of the route $r$\\
%         \hline
%         $P$ & Period\\
%         \hline
%         $\tau$ & Size of a message\\
%         \hline
%         $ [t(v,r)]$& Set of time slots used by route $r$ at vertex $v$ \\
%         & in a period $P$\\
%         \hline 
%         ${\cal M}=(m_0, \ldots ,m_{n-1})$& Assignment: an offset for each route)\\
%              \hline 
%         $w_r$& Waiting time of the route $r$\\
%            \hline 
%         $PT(r)$& Process time of the route $r$\\
%           \hline 
%         $MT({\cal M})$& Maximum process time of the assignment\\
%    \hline
%
%      \end{tabular}
%      \end{center}
%\section{Hardness of PRA}
%  \label{sec:complexity}
% In this section we always assume that the size of a message $\tau$ is equal to one. 
% We will prove the hardness of PRA and PALL for $\tau =1$ which implies the hardness of the general problems. 
%Consider an instance of the problem PRA, i.e., a routed network $(G,\cal{R})$ and a period $P$.
%The {\bf conflict depth} of a route is the number of arcs of the route which also belong to other routes.
%The conflict depth of a routed network $(G,\cal{R})$ is the maximum of the conflict depth of its routes.
%The {\bf load} of a routed network is the maximal number of routes sharing the same arc.
%Remark that a $(P,1)$-periodic assignment must satisfy that $P$ is larger or equal to the load.
%We give two alternate proofs that PRA is $\NP$-complete.
%The first proof works already for conflict depth two. Remark that for conflict depth one,
%the graph can be seen as a set of disjoint pair of routes, on which PRA and PALL can be solved in linear time. 
% The second proof reduces the problem to graph coloring and implies inapproximability when one tries to find the smallest possible $P$. \\
% Finally, it is easy to see that PRA is easy on trees and it may be interesting to study its complexity on 
% bounded treewidth networks, since it is a common property of real networks~\cite{de2011treewidth}.
% \begin{theorem}
%Problem PRA is $\NP$-complete on the class of routed networks with conflict depth two.
%\end{theorem}
% \begin{proof}
% Problem $PRA$ is in $\NP$ since given an offset for each route in an assignment, it is easy to check in linear time in the number of arcs whether there are collisions.
%  Let $H=(V,E)$ be an undirected graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is arc-colorable
%  with $d$ or $d+1$ colors. The arc coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. To do that, we define from $H$ a routed network $(G,{\cal R})$ as follows.
%  The vertices of $G$ are $v_1, v_2$ for each $v$ in $V$ and $s_{u,v}, t_{u,v}$ for each $(u,v) \in E$.
%  For each arc $(u,v) \in E$, there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$. 
%   To define ${\cal R}$ an arbitrary orientation of each arc $(u,v)$ is chosen. 
%   Then for each arc $(u,v)$ there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.
%  All these arcs are of weight $0$. The set of arcs of G is the union between all the arcs of the previously defined routes.
%  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $(d,1)$-periodic assignment
%  of $(G,{\cal R})$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its arcs by the integers
%  in $\{0,\dots,d-1\}$ and we have a bijection between $d$-colorings of $H$ and offsets of the routes of ${\cal R}$.
%  By construction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent arcs have the same color. Therefore we have reduced arc coloring to PRA by a polynomial time transformation which concludes the proof. 
% \end{proof}
% Remark that we have used zero weights in the proof. If we ask the weights to be strictly positive, which makes sense in our model since
%they represent the delay of physical links, it is easy to adapt the proof. We just have to set them so that in any route the weight at $u_1$ is equal to $d$ and thus equal to $0$ modulo $d$. We now lift this hardness result to the problem PALL.
%\begin{corollary}
%Problem PALL is $\NP$-complete on the class of routed networks of conflict depth two.
%\end{corollary}
% We consider $((G,{\cal R}),P,\tau)$ an instance of $PRA$. We assume that no vertex is the first of some route and the last of another one. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to this kind of instance $\NP$-complete. 
% Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We define $(G',{\cal R}')$ a symmetric routed network from $(G,{\cal R})$ where for every route we add a symmetric route with new arcs of opposite orientation and the same weights.
% We prove that the instance $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL if and only if $((G,{\cal R}),P,\tau)$ is in $PRA$.
% If $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL, then a $(P,\tau)$-assignment of $(G',{\cal R'})$ restricted to $\R$ is a $(P,\tau)$-assignment of $((G,{\cal R})$ since they cannot be any collision between routes of ${\cal R}$.
% Assume now that $((G,{\cal R}),P,\tau)$ is in $PRA$. First remark that the waiting time of each route is by definition less than $P$ and thus the maximal process time is always less than $T_{max}$, therefore the constraint on $T_{max}$ is satisfied. Moreover a $(P,\tau)$-assignment of $(G,{\cal R})$ can be extended into a $(P,\tau)$-assignment of $(G',{\cal R'})$ in the following way. For each route $r \in \cal{R}$ of last vertex $v$, the time at which the message arrives is $t(v,r)$, then we choose as offset for $\rho(r) \in \cal{R}'$ $-t(v,r) \mod P$. The symmetry ensures that each new route $\rho(r)$ in ${\cal R'}$ uses exactly the same times slot as $r$ one each of its node and thus avoid collisions.
%Let MIN-PRA be the problem, given a routed network and an assignment, to find the minimal period $P$ such that there is a $P$-periodic assignment. 
%\begin{theorem}\label{th:inapprox}
%If $\P \neq \NP$, problem MIN-PRA on the classes of routed networks of load two cannot be approximated in polynomial time within a factor $n^{1-o(1)}$ where $n$ is the number of routes.
%\end{theorem}
% We reduce graph coloring to PRA. Let $H$ be a graph instance of the $k$-coloring problem. 
% We define ${\cal R}$ in the following way: for each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$.
% Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. Note that it is easy to build a graph with such routes as in Fig.~\ref{fig:reduction}.
% Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
% by converting an offset of a route into a color of a vertex and reciprocally. Therefore if we can approximate the minimum value of $P$ within some factor, we could approximate the minimal number of colors needed to color a graph within the same factor. The proof follows from the hardness of approximability of finding a minimal coloring~\cite{zuckerman2006linear}.
%In particular, this reduction shows that even with small maximal load, the minimal period can be large.
%    \hspace{0.5cm}\scalebox{0.35}{
%    \begin{tikzpicture}
%      \node[draw,circle] (t3) at (12, 3) {$t_2$}; 
%      \tikzstyle{VertexStyle}=[shape = circle, draw, minimum size = 20pt]
%      VertexStyle/.append style = {blue} }
%	\Vertex[x=-8,y=3]{1}
%      VertexStyle/.append style = {green} }
%	  \Vertex[x=-7,y=5]{2}
%      VertexStyle/.append style = {red} }
%	  \Vertex[x=-6,y=4]{3}
%      VertexStyle/.append style = {black} }
%      \Edge(s2)(A)
%      \Edge(A)(B)
%      \Edge(B)(C)
%      \Edge(C)(D)
%      \Edge(D)(t2)
%      \Edge(s3)(E)
%      \Edge(E)(F)
%      \Edge(F)(t3) 
%      \Edge(s1)(A)
%      \Edge(A)(B)
%      \Edge(B)(E)
%      \Edge(E)(F)
%      \Edge(F)(t1)
%      EdgeStyle/.append style = {black,-} }
%      \Edge(1)(2)
%      \Edge(1)(3)
%    \node (1) at (-3,4){\Huge $\rightarrow$};
%%     
%%     \node (2) at (-7,0){\Huge H};
%%     \node (3) at (10,0){\Huge G};
%    \end{tikzpicture}
%    }
%    \caption{Reduction from k-coloring to MIN-PRA}
%    \label{fig:reduction}

\section{Solving \pazl on Star Routed Networks} \label{sec:PAZL}

  In this section, we deal with the problem \pazl on star routed networks. Since we want an assignment with waiting times zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. Moreover choosing the offset $m_r$ of a forward route $r$ also sets the offset of the backward route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.  Consider an assignment $m$ of a star routed network and let $m'_r= m_{r} - t(c_s,r) \mod P$.
  Then $m'$ is an assignment of the same star routed network where, for all $i$, the weights of the arcs from $s_i$ to $c_s$ are set to $0$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a star routed network, and let $n\tau + 2(\lambda(r_{n-1}) - \lambda(r_{0})) \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which is given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{r_i} = i\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

@@

    \todo{Introduire la load là et faire un théorème avec ce paramètre ?}

    We propose a greedy algorithm to build a $(P,\tau)$-periodic assignment, which provably works when
    the period is large enough with regard to the number of routes and size of messages. In other words, 
    we can always find a solution when the network is far from saturated. This algorithm defines an upper bound for the period. As a reminder, the lower bound is $n\tau$, that is, the size taken by the messages to be emitted.

    \begin{proposition}
     Let $(G, {\cal R})$ be a star routed network, and let $ 3n\tau \leq P$, then there is $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with all waiting times $0$ which can be found in time $O(n^2)$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive intervals of size $\tau$ that we call macro-slots. The algorithm works by choosing an offset for each route in the following way: try all offsets which put the message in a yet not used macro-slot in the forward
     period. Since the choice of an offset also fixes the position of the message in the backward period, chose the first one which does not create a collision. We now prove that this algorithm always finds a $(P,\tau)$-periodic assignment without waiting time when $P \geq 3n\tau$.

@@

\begin{theorem}\label{th:FPT}
$\pall \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $\mathcal{M}$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $\mathcal{M}$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R}$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to 
not use route $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select. The value of the offset of the selected route is entirely determined by the values of the offsets of routes in $S$. Therefore there are at most $n!$ different compact assignments with offset $r_0$ fixed to $0$.

@@

    \subsection{Experimental evaluation}\label{sec:exp_PAZL}
      The following experimental results compare the three presented algorithms.
      Notice that both Greedy algorithm and Shortest-Longest are polynomial time algorithm but are not always able to find a solution, depending on $P$, $\tau$, $n$ and the size of the routes. On the other hand, the exhaustive search finds an optimal solution if it exists, but works in exponential time. We compare the performance of the algorithms in two different regimes: routes are either short with regard to $\tau$, or unrestricted.
      From our C-RAN context we choose the following parameters: the number of routes 
      is at most $n = 20$, $\tau$ is equal to $2500$. Moreover the period is always larger than $n\tau$ otherwise no assignment is possible and smaller than $3n\tau$ otherwise the greedy algorithm always returns a solution.

@@

      We define the load of the network as $\frac{n\tau}{P}$ (here a value between one and three),
      it is the proportion of time slots used by messages on the central arc. In our experiments we will try to understand how our algorithms work with regards to the load. To change the load, we fix the parameters $\tau$ and $n$ and modify the period $P$, which allows for a smoother control of the load and does not change the run time of our algorithms.

@@

      First we consider routes which are shorter than $\tau$, that is a message cannot be contained completely in a single arc which is very common for our application. We generate star routed networks in which the weights of the arcs $(c_t,t_i)$ are drawn uniformly between $0$ and $700$ slots. It corresponds to messages of approximately $1$Mb, links of bandwidth $10$Gbps and length less than $5$km between the BBU and the RRH.

      Our aim is to understand how well the algorithms are working under high load. To do that we try to evaluate the minimal period (that is the highest possible load) for which a $(P,\tau)$-periodic assignment can be found by each algorithm. 
%       To that end we use a linear search on $P$, since a dichotomy method would not work because of Lemma~\ref{lemma:monotonic}.

      In our experiment, we generate $1000$ random instances of \pazl for $1$ to $12$ routes. 
      We measure the average minimal period, for each algorithm. Remind that the exhaustive search always finds the best possible period for a given instance. The lower and upper bound $n\tau$ and $3n\tau$ are also represented.
      \todo{présenter ça comme une average load ?}
      \begin{figure}[h]
      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.5\textwidth]{periode_petite.pdf}
      \end{center}
            \vspace{-0.75cm}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}
      First, we remark that the period found by the exhaustive search is only very slightly above the lower bound of $n\tau$, which means that, in this regime, it is very well justified to look for a solution without waiting time even for a highly loaded network.

@@

      We now want to understand the performance of these algorithms when the size of the routes is unbounded. In this experiment we fix the number of routes to $8$ and the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots (in the same range as the period). We measure for each algorithm its percentage of success, for periods from $20000$ to $50000$ which corresponds to a load of $100\%$ down to $40\%$.

\begin{figure}[H]

@@

      In this regime, the performances of Shortest-Longest are abysmal since it depends on the difference between the longest and the smallest route which is large here. On the other hand, the greedy algorithm has a performance not so far from the case of short routes, which is expected since it does not directly depend on the size of the route. When repeating the experiments of the previous section on routes of arbitrary length, we find that on average the greedy algorithm finds a solution if $P>1.89n\tau$.

      When the load is larger than $50\%$ (period less than $40000$), the exhaustive search finds  more solutions than the greedy algorithms which justifies its use. However, for load larger than $80\%$ (period less than $25000$) there are many instances for which there are no solutions to \pazl.
      It means that with long routes and high load, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general \pall problem in our next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignment without waiting times.

      The computation time of the exhaustive search is in $O(n!)$, hence we need to study its scalability when $n$ grows. In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots. We fixed a period of $1.05n\tau$, which corresponds to $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks. This shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but on more routes we should improve it further.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $11$\\
    \hline
      \end{tabularx}
      \end{center}

@@

\begin{figure}[H]

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[H] 
          \includegraphics[width=0.4\textwidth]{departs_gp_25000.pdf}
          \vspace{-0.5cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

>>>>

\begin{abstract}
A recent trend in mobile networks is to centralize in distant data-centers the processing units which were attached to 
antennas. The main challenge is to guarantee that the latency of the periodic messages sent from the antennas to their processing
units and back fulfills protocol time constaints. We show that traditional statistical multiplexing does not allow such a low latency, due to collisions and buffering at nodes. Hence, we propose in this article to use a deterministic scheme for sending periodic messages without collisions in the network thus without buffering and the resulting additional latency. 
We give several algorithms to compute such schemes for a common topology where one link is shared by all antennas.
We show that there is always a solution when the routes are short or the load is small. When the parameters are unconstrained,
and some buffering is allowed in the processing units, we propose an algorithm (PMLS) adapted from a classical scheduling method.
The experimental results show that even under full load, most of the time PMLS finds a deterministic sending scheme with no latency.
\end{abstract}

@@

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN, or Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost ownership. The main challenge for this type of architecture is the latency in the transfer process that is an important factor for dimensioning the network. The latency is measured between the sending of a message by the Remote Radio Heads (RRHs) on the field and the receptions of the answers, computed by real-time virtualized network functions on Baseband Units (BBUs), in the cloud. For example, even if next radio protocols are less stringent, some standards requires meeting time constraints for functions like HARQ (Hybrid Automatic Repeat reQuest) that needs to be processed in $3$ms~\cite{bouguen2012lte}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer between RRHs and BBUs.

@@

%       In Sec.~\ref{sec:\pall} we will study algorithms which solve the search version of \pall (computing an assignment), also denoted by \pall for simplicity. \todo{phrase utile finalement ?} 
      It turns out that the problem \pall is hard to solve for \emph{general} routed networks. The long version of this article contains complete proofs of the $\NP$-hardness or hardness of approximation of \pall and of several of its variants.

@@

       We consider a simpler version of \pall, that we call {\bf Periodic Assignment for Zero Latency} or \pazl: we ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$}. We introduce \pazl because it is simpler to study and we are able to prove theoretical results and find better algorithms than for \pall. Moreover, as we  experimentally show in Sec.~\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to \pazl is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.

@@

      Let us define a family of simple routed networks which corresponds to a Point-to-Multipoint fronthaul (PtMP), intended for C-RAN \cite{tayq2017real}. 
      The graph $G$ has two sets of vertices, $S=\{s_0,...,s_{n-1}\}$ and $T=\{t_0,...,t_{n-1}\}$ of cardinality $n$ and two special nodes, the central source node $c_s$ and the central target node $c_t$.
      There is an arc between $c_s$ and  $c_t$ called the central arc and for all $i<n$ there is an arc between $s_i$ and $c_s$ and between $t_i$ and $c_t$. All the symmetric arcs are also in the graph with the same weights.
      The forward routes are the directed paths $r_i = (s_i,c_s,c_t,t_i)$ and the backward routes are $\rho(r_i) =(t_i,c_t,c_s,s_i)$. The routed network $(G, \{r_i,\rho(r_i)\}_{i<n})$ is called a \textbf{star routed network}.

@@

@@

  Star routed networks seem really simple, but \pall or \pazl may be $\NP$-hard restricted to these networks.
  We conjecture it is the case since  a variant of \pazl where the central link is unidirectional (collisions can happen between forward and backward routes) can be shown to be $\NP$-complete. It is done by a reduction from the subset sum problem as for a similar problem of scheduling pair of tasks~\cite{orman1997complexity}.

@@

         \section{Solving \pazl on Star Routed Networks}\label{sec:PAZL}

       In this section, we deal with the problem PAZL on star routed networks. Since we want an assignment with waiting times zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. Moreover choosing the offset $m_r$ of a forward route $r$ also sets the offset of the backward route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.  Consider an assignment $m$ of a star routed network and let $m'_r= m_{r} - t(c_s,r) \mod P$. Then $m'$ is an assignment of the same star routed network where, for all $i$, the weights of the arcs $(s_i,c_s)$ are set to $0$. Therefore we assume from now on that the \emph{weight of the arcs $(s_i,c_s)$ are all equal to $0$}.

@@

      \begin{proposition} Let $(G, {\cal R})$ be a star routed network, and let $n\tau + 2(\lambda(r_{n-1}) - \lambda(r_{0})) \leq P$. There is a $(P,\tau)$-periodic assignment of $(G, {\cal R})$ with waiting times $0$ given by Shortest-Longest in time $O(n\log(n))$.\label{prop:SL}
      \end{proposition}
      \begin{proof}
       Since $m_{r_i} = i\tau$, $[t(c_s,r_{i})] = \{i\tau,\dots, (i+1)\tau -1\}$ and there are no collision on the forward period.

@@

      We define the \textbf{load} of a star routed network as $\frac{n\tau}{P}$, it is the proportion of time slots used by messages on the central arc in a period. Therefore if the load is larger than $1$ there cannot be an assignment. We propose a greedy algorithm to build a $(P,\tau)$-periodic assignment, which always finds an assignment when the load is less than $1/3$. Therefore in the rest of the article we will be only concerned with load larger than $1/3$.

    \begin{proposition}
    There is a $(P,\tau)$-periodic assignment of a star routed network with waiting times $0$ if the load is less than $1/3$ and it can be found in time $O(n^2)$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive intervals of size $\tau$ that we call macro-slots. The algorithm works by choosing an offset for each route in the following way: try all offsets which put the message in a yet not used macro-slot in the forward
     period. Since the choice of an offset also fixes the position of the message in the backward period, chose the first one which does not create a collision. We now prove that this algorithm always finds a $(P,\tau)$-periodic assignment without waiting time when $P \geq 3n\tau$ that is the load is less than $1/3$.

@@

\begin{theorem}\label{th:FPT}
$\pall \in \FPT$ when parametrized by the number of routes.
\end{theorem}
\begin{proof}
Let $(G, {\cal R})$ be a star routed network and let $m$ be a $(P,\tau)$-periodic assignment of $(G, {\cal R})$. First remark that for a given assignment and a route $r_0$ with offset $m$, by removing $m$ to all offsets, we can always assume that its offset is zero. Therefore we need only to consider all \emph{compact assignments} with an \emph{offset $0$} for the route $r_0$. 
We now evaluate the number of compact assignments and prove that it only depends
on $n$ the number of routes which proves the theorem. 
We give a way to build a compact assignment $m$ by determining its offsets one after the other. We fix an arbitrary total order on ${\cal R}$.
First a route $r_0$ is chosen arbitrarily and its offset set to $0$. 
Then at each step, if the offsets of $S \subseteq  {\cal R}$ have been chosen,
we select the smallest route $r$ in $S$ for the order. 
Then we select a route in $r' \in {\cal R} \setminus S$ and set its offset such that 
if we remove $1$ then $r'$ collides with $r$. Note that if $r$ is a forward route (resp. a backward route) then $r'$ is also a forward route (resp. a backward route). We can also decide to definitly skip $r$. At a given step of the algorithm, if $|S| = 2i$, we have $n-i$ choices 
of routes to select. The value of the offset of the selected route is entirely determined by the values of the offsets of routes in $S$. Therefore there are at most $n!$ different compact assignments with offset $r_0$ fixed to $0$.

@@

   \subsection{Experimental evaluation}\label{sec:exp_PAZL}
      The following experimental results compare the three presented algorithms.
      Notice that both Greedy algorithm and Shortest-Longest are polynomial time algorithms but are not always able to find a solution, depending on the load or the size of the routes. On the other hand, the exhaustive search finds an optimal solution if it exists, but works in exponential time. We compare the performance of the algorithms in two different regimes: routes are either short with regard to $\tau$, or unrestricted.
      From our C-RAN context we choose the following parameters: the number of routes is at most $n = 20$, $\tau$ is equal to $2500$.
%       Moreover the period is always larger than $n\tau$ otherwise no assignment is possible and smaller than $3n\tau$ otherwise the greedy algorithm always returns a solution.

@@

       In our experiments we try to understand how our algorithms work with regards to the load. To change the load, we fix the parameters $\tau$ and $n$ and modify the period $P$, which allows for a smoother control of the load and does not change the run time of our algorithms.

@@

      First we consider routes which are shorter than $\tau$: a message cannot be contained completely in a single arc which is very common in our applications. We generate star routed networks in which the weights of the arcs $(c_t,t_i)$ are drawn uniformly between $0$ and $700$. It corresponds to messages of approximately $1$Mb, links of bandwidth $10$Gbps and length of routes less than $5$km between the BBU and the RRH.

      Our aim is to understand how well the algorithms are working under high load. To do that we evaluate the highest load 
      under which a $(P,\tau)$-periodic assignment can be found by each algorithm when we change the number of routes. 
      In our experiment, we generate $1000$ random instances of \pazl for $1$ to $14$ routes. We measure the average maximal possible load, for which each algorithm finds a solution. Remind that the exhaustive search always finds the highest load for a given instance. 
%       The lower and upper bound $n\tau$ and $3n\tau$ are also represented.

      \begin{figure}[h]

      \begin{center}
	 \hspace{-0.5cm}\includegraphics[width=0.5\textwidth]{periode_petite.pdf}
      \end{center}
            \vspace{-0.75cm}
      \caption{Minimum period averaged over $1000$ random instances}
      \end{figure}
      First, we remark that the period found by the exhaustive search is only very slightly above the lower bound of $n\tau$, which means that, in this regime, it is very well justified to look for a solution without waiting time even for a highly loaded network.

@@

      We now want to understand the performance of these algorithms when the size of the routes is unbounded. In this experiment we fix the number of routes to $8$ and the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots (in the same range as the period). We measure for each algorithm its percentage of success, for load of $100\%$ down to $40\%$.

\begin{figure}[h]

@@

      In this regime, the performances of Shortest-Longest are abysmal since it depends on the difference between the longest and the smallest route which is large here. On the other hand, the greedy algorithm has a performance not so far from the case of short routes, which is expected since it does not directly depend on the size of the route. In fact, if we do the previous experiment 
      for short routes but with long routes, we find that, on average, the greedy algorithm finds a solution when the load is less than $53\%$.

      When the load is larger than $50\%$, the exhaustive search finds more solutions than the greedy algorithms which justifies its use. However, for load larger than $80\%$ there are many instances for which there are no solutions to \pazl.
      It means that with long routes and high load, looking for an assignment without waiting time is far too restrictive. That is why we present algorithms for the general \pall problem in our next section. We will test them on $8$ long routes and a load between $100\%$ and $80\%$ for which, as shown here, there are often no assignment without waiting times.

      The computation time of the exhaustive search is in $O(n!)$ as shown int Th.~\ref{th:FPT}, hence we need to study its scalability when $n$ grows. In the following experiment, the weights of the arcs $(c_t,t_i)$ are drawn following an uniform distribution between $0$ and $30000$ slots. We chosed $95\%$ of load.
%       , that corresponds to loads of $\simeq 70\%$ (for $1$ route) to $\simeq 98\%$ (for 21 routes).
      The following table shows the time before the exhaustive search ends, for $8$ to $16$ routes, averaged on $100$ random star routed networks. This shows that for less than $15$ routes, which corresponds to all current topologies, the algorithm is efficient enough, but on more routes we should improve it further.
         \begin{center}
         \begin{tabularx}{0.5\textwidth}{|l|X|X|X|X|}
    \hline
   $n$ & $8$ & $10$& $12$& $16$\\
    \hline
   Time (s) & $6.10^{-5}$&$8.10^{-4}$&$2.10^{-2}$& $11$\\
    \hline
      \end{tabularx}
      \end{center}

@@

\begin{figure}[h]

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}
\vspace{-0.75cm}
      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
          \vspace{-1cm}
\begin{figure}[h] 
          \includegraphics[width=0.4\textwidth]{departs_gp_25000.pdf}
          \vspace{-0.5cm}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}
